---
title: "Teen market segments via k-means"
theme: paper
output:
  html_document:
    theme: default
---

```{r}
teens <- read.csv('snsdata.csv')
```

Let us first see the gender distribution of our dataset. 
(NAs are excluded in the result)
```{r}
table(teens$gender)
```

The age distribution within the dataset.
```{r}
summary(teens$age)
```

The min age is 3 an the max age is 106 which means we need to clean up 
the age. 
```{r}
teens$age <- ifelse(teens$age >= 13 & teens$age < 20, teens$age, NA)
```

Dummy coding of gender.
```{r}
teens$female <- ifelse(teens$gender == 'F' & !is.na(teens$gender), 1, 0)
teens$no_gender <- ifelse(is.na(teens$gender), 1, 0)
```

Check the gender again.
```{r}
table(teens$gender)
table(teens$female)
table(teens$no_gender)
```

**Dealing with missing ages**

The mean age is
```{r}
mean(teens$age, na.rm = T)
```

Age by graduation year is
```{r}
aggregate(data = teens, age ~ gradyear, mean, na.rm = T)
```

Thus the age differs by one year per change in graduation year. 
```{r}
ave_age <- ave(teens$age, teens$gradyear, FUN = function(x) mean(x, na.rm = T))
teens$age <- ifelse(is.na(teens$age), ave_age, teens$age)
summary(teens$age)
```

**Training the model**
```{r}
features <- teens[5:40]
# z-score normalization
features.z <- as.data.frame(lapply(features, scale))
```

Let us first try to divide the data into 5 clusters.
```{r}
teen_clusters <- kmeans(features.z, 5)
```

The sizes of the clusters are 
```{r}
teen_clusters$size
```

The centers of the clusters are 
```{r}
teen_clusters$centers[, 1:8]
```

Set the cluster as an extra column of the dataset.
```{r}
teens$cluster <- teen_clusters$cluster
```

Let us try to visualize the clusters.
```{r}
plot(c(0, 10), c(0, 10), type = 'n', 
     xlab = 'one feature', ylab = 'another feature')
points(x = jitter(subset(teens, cluster == 1)$rock, 1), 
     y = jitter(subset(teens, cluster == 1)$mall, 1),
     pch = 1, col = rgb(red = 1, green = 0, blue = 0, alpha = 0.5))
points(x = jitter(subset(teens, cluster == 2)$rock, 1), 
     y = jitter(subset(teens, cluster == 2)$mall, 1),
     pch = 1, col = rgb(red = 0, green = 1, blue = 0, alpha = 0.5))
points(x = jitter(subset(teens, cluster == 3)$rock, 1), 
     y = jitter(subset(teens, cluster == 3)$mall, 1),
     pch = 1, col = rgb(red = 0, green = 0, blue = 1, alpha = 0.5))
```

The results are not pretty because two many features were used for 
the clustering while the result is only plotted in 2 dimension.

Now let us take a look at the scatterplot matrices. It is obvious from the
plots that most data points are concentrated near the origin/axis and this 
makes it harder to draw the boundaries between different clusters. As a 
consequence, it is questionable whether the data points form clusters or not.

```{r}
pairs(features.z[1:5])
```