---
title: "Diagnosing breast cancer with kNN"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load data from csv file. The data file is downloaded from https://www.kaggle.com/uciml/breast-cancer-wisconsin-data/version/2#data.csv

```{r}
data <- read.csv('data.csv', stringsAsFactors = F)
```

View structure of data set
```{r}
str(data)
```

Drop id from the data
```{r}
data <- data[-1]
```

Let us see the distribution of diagnosis first
```{r}
table(data$diagnosis)
```

Convert diagnosis into factors
```{r}
data$diagnosis <- factor(data$diagnosis, levels = c('M', 'B'), labels = c('Malignant', 'Benign'))
```

Normalizing the numerical features
```{r}
normalize <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

data.n <- as.data.frame(lapply(data[2:31], normalize))
summary(data.n)
```

We have 569 observations in total. The first 469 observation serves as our training set while the rest 100 will be our
test data.
```{r}
data.train <- data.n[1:469, ]
data.test <- data.n[470:569, ]
```

Store the labels for the dataset.
```{r}
train.labels <- data[1:469, 1]
test.labels <- data[470:569, 1]
```

Install and load the package for classification
```{r}
#install.packages('class')
library(class)
```

Make prediction with kNN algorithm
```{r}
test.pred <- knn(train = data.train, test = data.test, cl = train.labels, k = 3)
```

Evaluate the performance of our model
```{r}
# install.packages('gmodels')
library(gmodels)
CrossTable(x = test.labels, y = test.pred)
```

In the cross table, there are four kinds of outcomes:

* True positive
* True negative
* False positive (Bad)
* False negative (Worse)
